---
title: "Assignment 4 - Coordinating Heart Rate"
author: "Riccardo Fusaroli"
date: "November 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
install.packages("knitr")
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

```

## Analysing Heart Rate and Respiration data

The goal of this assignment is to first familiarize you with heart rate, and respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.
- Artefacts 
- Down sample
- Scale-makes the signals more comparabel - 1) subtract the mean (so both signals start at 0), 2) ???
- PLOT THE DATA! - save the plot - ggsave()
- run crqa optimize param 
- Do this for all files in a loop

Second loop: 
- Choose one embed, radius and delay parameter (use media) - for HR radius add 1.2 to the mean - should give a mean REC of 4.5 - makes sure that all the data are adequately represented in the analysis 
- Run crqa on all the data files with these parameters 

```{r}
setwd("~/Documents/Experimental Methods III/Assignment_4/Data")
install.packages("devtools")
devtools::install_github("LudvigOlsen/groupdata2", force=TRUE)
install.packages("pacman")
library(pacman)
p_load(dplyr, groupdata2, stringr, ggplot2, tidyverse, rgl, tseriesChaos, SparseM, crqa)

#Function for removing outliers (artefacts)
removeOuts = function(ts,threshold){ts[ts > (mean(ts,na.rm=T) +           (threshold*sd(ts,na.rm=T))) |ts < (mean(ts,na.rm=T) -           (threshold*sd(ts,na.rm=T)))] = mean(ts,na.rm=T)  
return(ts)}

#Loop doing pre-processing 
filelist = list.files("~/Documents/Experimental Methods III/Assignment_4/Data", pattern = ".*.csv")
setwd("~/Documents/Experimental Methods III/Assignment_4/Data")

for (i in filelist){
  data = read.delim(i, header=TRUE, sep = ",")
  
    #removing outliers 
  data$HR1 = removeOuts(data$HR1, 2.5) #Using threshold of 2.5 standard deviations
  data$HR2 = removeOuts(data$HR2, 2.5)
  data$Resp1 = removeOuts(data$Resp1, 2.5)
  data$Resp2 = removeOuts(data$Resp2, 2.5)
  
  #downsample
  data = data %>% 
  group(n = 100, method = 'greedy') %>%    
  dplyr::summarise(time = mean(time,na.rm=T),HR1 = mean(HR1,na.rm=T),
                   HR2 = mean(HR2,na.rm=T),Resp1 = mean(Resp1,na.rm=T),
                   Resp2 = mean(Resp2,na.rm=T))
  
  
  #scaling 
  data$Resp1S=scale(data$Resp1) 
  data$Resp2S=scale(data$Resp2)  
  data$HR1S=scale(data$HR1)
  data$HR2S=scale(data$HR2)
  
  #Save the new data with individual names for each file/loop
  file1 = paste0("pp_",i )
  write.csv(data, file.path("prepro_data", file1))
  
}

#PLOT TIMESERIES
filelist = list.files("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data", pattern = "^pp_Study.*.csv")

setwd("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data")
for (i in filelist){
  data = read.delim(i, header=TRUE, sep = ",")
  
  #Extract group information 
  group = str_extract(i, regex("G\\d{1}"))
  #Extract trial information 
  trial = str_extract(i, regex("T\\d{1}"))
  
  plot_HR = ggplot(data, aes(time, HR1S), colour="darkgreen") +
    geom_line()+
    geom_line(aes(time, HR2S), colour = "darkred") +
    labs(x = "time", y="HR", header = "")
  
  ggsave(paste0("HR_plot",group,trial ), device = "png",plot=plot_HR)
  
  plot_Resp = ggplot(data, aes(time, Resp1S), colour="darkgreen") +
    geom_line()+
    geom_line(aes(time, Resp2S), colour = "darkred") +
    labs(x = "time", y="HR", header = "")
  
  ggsave(paste0("Resp_plot",group,trial), device ="png",plot=plot_Resp)
 }

#OUTPUT OF PLOTS 
#We exclude group 2, trial 1 and 2 RESP data - looks like something was wrong with the measuring device (the belt)


pp_filelist = list.files("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data", pattern = ".*.csv")

HR_Embed = NULL 
HR_Delay = NULL 
HR_Threshold = NULL 

Resp_Embed = NULL 
Resp_Delay = NULL 
Resp_Threshold = NULL 

n = 1
setwd("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data")
for (i in pp_filelist){
  data = read.delim(i, header=TRUE, sep = ",")
  
  #Run optimize param and store in df 
  par = list(lgM =  15, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample   = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw =   0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
  
  ans_HR = try(optimizeParam(data$HR1S, data$HR2S, par, min.rec = 2, max.rec = 8))
  
  ans_Resp = try(optimizeParam(data$Resp1S, data$Resp2S, par, min.rec = 2, max.rec = 8))
  
  if (length(ans_HR)<2){
  HR_Delay[n] = NA
  HR_Embed[n] = NA
  HR_Threshold[n] = NA
  } else {
  HR_Delay[n] = ans_HR$delay
  HR_Embed[n] = ans_HR$emddim
  HR_Threshold[n] = ans_HR$radius
  }
  
  if (length(ans_Resp)<2){
  Resp_Delay[n] = NA
  Resp_Embed[n] = NA
  Resp_Threshold[n] = NA
  } else {
  Resp_Delay[n] = ans_Resp$delay
  Resp_Embed[n] = ans_Resp$emddim
  Resp_Threshold[n] = ans_Resp$radius
  
  n=n+1
  
  }
  }


crqa_df = data.frame(HR_Delay, HR_Embed, HR_Threshold, Resp_Delay, Resp_Embed, Resp_Threshold)

#Safe crqa parameter df  
write.csv(crqa_df, file="crqa_param_df")






#RUN LOOP WITH fitCrqa function (from Riccardo's slides)
#Extract study, group, trial, condition
crqa_filelist = list.files("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data", pattern = "^pp_Study.*.csv")

#RUN THE CRQA ANALYSIS ON THE DATAFILES 

HR_Delay = median(crqa_df$HR_Delay, na.rm=TRUE)
HR_Embed = median(crqa_df$HR_Embed, na.rm=TRUE)
HR_Radius = median(crqa_df$HR_Threshold, na.rm=TRUE)+1.2

Resp_Delay = median(crqa_df$Resp_Delay, na.rm=TRUE)
Resp_Embed = median(crqa_df$Resp_Embed, na.rm=TRUE)
Resp_Radius = median(crqa_df$Resp_Threshold, na.rm = TRUE)

study = NULL
group = NULL 
trial = NULL 
condition = NULL 
HR_RR = NULL 
HR_L = NULL 
HR_DET = NULL 
HR_maxL = NULL 
HR_TT = NULL 
HR_LAM = NULL 
HR_ENTR = NULL
Resp_RR = NULL 
Resp_L = NULL 
Resp_DET = NULL 
Resp_maxL = NULL 
Resp_TT = NULL 
Resp_LAM = NULL 
Resp_ENTR = NULL
Type= NULL 

setwd("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data")
n=1
for (i in crqa_filelist){
  ts = read.delim(i, header=TRUE, sep=",")
  
  #Extract study information 
  study[n] = str_extract(i, regex("\\d{1}"))
  #Extract group information 
  group[n] = str_extract(i, regex("G\\d{1}"))
  #Extract trial information 
  trial[n] = str_extract(i, regex("T\\d{1}"))
  #Extract condition information 
  condition[n] = str_extract(i, regex("[A-Za-z]{6,}"))
  
  Type[n] = "real"
  
  crqa_HR = try(crqa(ts$HR1S, ts$HR2S, delay=HR_Delay, embed=HR_Embed, radius=HR_Radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  
  HR_RR[n] = crqa_HR$RR
  HR_L[n] = crqa_HR$L
  HR_DET[n] = crqa_HR$DET
  HR_maxL[n] = crqa_HR$maxL
  HR_TT[n] = crqa_HR$TT
  HR_LAM[n] = crqa_HR$LAM
  HR_ENTR[n] = crqa_HR$ENTR
  
  crqa_Resp = try(crqa(ts$Resp1S, ts$Resp2S, delay=Resp_Delay, embed=Resp_Embed, radius=Resp_Radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  Resp_RR[n] = crqa_Resp$RR
  Resp_L[n] = crqa_Resp$L
  Resp_DET[n] = crqa_Resp$DET
  Resp_maxL[n] = crqa_Resp$maxL
  Resp_TT[n] = crqa_Resp$TT
  Resp_LAM[n] = crqa_Resp$LAM
  Resp_ENTR[n] = crqa_Resp$ENTR

  n=n+1
  }

crqa_data = data.frame(study, trial, group, condition, HR_RR, HR_L, HR_DET, HR_maxL, HR_TT, HR_LAM, HR_ENTR, Resp_RR, Resp_L, Resp_DET, Resp_maxL, Resp_TT, Resp_LAM, Resp_ENTR, Type)

#Save the dataframe of all crqa measures for each file (in one df)
write.csv(data, file="crqa_data")

```


```{r}
#SHUFFLE THE DATA AND RUN CRQA ON THOSE - Store in a df
library(base)

setwd("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data")

#Shuffle data 
for (i in pp_filelist){
  basedata = read.delim(pp_filelist[1], header=TRUE, sep=",")
  shuff_data = basedata
  shuff_data$HR1S = sample(basedata$HR1S, replace=FALSE)
  shuff_data$HR2S = sample(basedata$HR2S, replace=FALSE)
  shuff_data$Resp1S = sample(basedata$Resp1S, replace=FALSE)
  shuff_data$Resp2S = sample(basedata$Resp2S, replace=FALSE)
  
  #Save the new data with individual names for each file/loop
  file1 = paste0("shuffled_",i )
  write.csv(shuff_data, file = file1)
}

 
shuff_filelist = list.files("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data", pattern = "shuffled.*.csv")
shuff_filelist

crqa_param_df <- read.csv("~/Documents/Experimental Methods III/Assignment_4/crqa_param_df")

#Run CRQA analysis on shuffled data 
HR_Delay = median(crqa_df$HR_Delay, na.rm=TRUE)
HR_Embed = median(crqa_df$HR_Embed, na.rm=TRUE)
HR_Radius = median(crqa_df$HR_Threshold, na.rm=TRUE)+1.2

Resp_Delay = median(crqa_df$Resp_Delay, na.rm=TRUE)
Resp_Embed = median(crqa_df$Resp_Embed, na.rm=TRUE)
Resp_Radius = median(crqa_df$Resp_Threshold, na.rm = TRUE)

study = NULL
group = NULL 
trial = NULL 
condition = NULL 
HR_RR = NULL 
HR_L = NULL 
HR_DET = NULL 
HR_maxL = NULL 
HR_TT = NULL 
HR_LAM = NULL 
HR_ENTR = NULL
Resp_RR = NULL 
Resp_L = NULL 
Resp_DET = NULL 
Resp_maxL = NULL 
Resp_TT = NULL 
Resp_LAM = NULL 
Resp_ENTR = NULL
Type= NULL 

setwd("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data")

n=1
for (i in shuff_filelist){
  ts = read.delim(i, header=TRUE, sep=",")
  
  #Extract study information 
  study[n] = str_extract(i, regex("\\d{1}"))
  #Extract group information 
  group[n] = str_extract(i, regex("G\\d{1}"))
  #Extract trial information 
  trial[n] = str_extract(i, regex("T\\d{1}"))
  #Extract condition information 
  condition[n] = str_extract(i, regex("Synchronous|TurnTaking|SelfPaced|Conversation"))
  
  Type[n] = "shuffled"
  
  crqa_HR = try(crqa(ts$HR1S, ts$HR2S, delay=HR_Delay, embed=HR_Embed, radius=HR_Radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  
  HR_RR[n] = crqa_HR$RR
  HR_L[n] = crqa_HR$L
  HR_DET[n] = crqa_HR$DET
  HR_maxL[n] = crqa_HR$maxL
  HR_TT[n] = crqa_HR$TT
  HR_LAM[n] = crqa_HR$LAM
  HR_ENTR[n] = crqa_HR$ENTR
  
  crqa_Resp = try(crqa(ts$Resp1S, ts$Resp2S, delay=Resp_Delay, embed=Resp_Embed, radius=Resp_Radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  
  
  Resp_RR[n] = crqa_Resp$RR
  Resp_L[n] = crqa_Resp$L
  Resp_DET[n] = crqa_Resp$DET
  Resp_maxL[n] = crqa_Resp$maxL
  Resp_TT[n] = crqa_Resp$TT
  Resp_LAM[n] = crqa_Resp$LAM
  Resp_ENTR[n] = crqa_Resp$ENTR
  n=n+1
  }

shuff_crqa_data = data.frame(study, trial, group, condition,HR_RR, HR_L, HR_DET, HR_maxL, HR_TT, HR_LAM, HR_ENTR, Resp_RR, Resp_L, Resp_DET, Resp_maxL, Resp_TT, Resp_LAM, Resp_ENTR, Type)

shuff_crqa_data = shuff_crqa_data[!shuff_crqa_data$condition=="SelfPaced",]

#Save the dataframe of all crqa measures for each file (in one df)
write.csv(shuff_crqa_data, file="shuff_crqa_data")

```




```{r}
#CREATING SURROGATE PAIRS AND RUN THE CRQA ON THESE - Store in a df

setwd("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data")

prepro_filelist = list.files("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data", pattern = "^pp_Study1.*.csv")

num=str_extract_all(prepro_filelist, "\\d+") #extract all numbers in the file name (i.e Study?_G?_T?)

df= as.data.frame(t(as.data.frame(num))) #create a transposed (i.e. "tipped") dataframe of these number
rownames(df)<-NULL #Remove rownames.

pairs=as.numeric(unique(df$V2)) #create pairs, i.e. unique group numbers

SurrogateList <- expand.grid(a = pairs, b = pairs) # expand.grid combines all possible combinations of a given vector
SurrogateList = subset(SurrogateList, a != b) # Ensure that pair 1 (i.e group1) doesn't get paired with it self.

setwd("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data")
# Per each condition
for (co in c("Synchronous","TurnTaking")){ #loop through a list of condintions 
  for (i in 1:nrow(SurrogateList)){ #and loop through all the pairings of the surrogate list
    
  #start recreate filenames from the working directory based on the matching from the surrogatelist
  f1=paste0('Study1_G',SurrogateList$a[i]) #Insert group numeber from Surrogatelist 
  f2=paste0('Study1_G',SurrogateList$b[i])
  
  file1=prepro_filelist[grepl(f1,prepro_filelist) & grepl(co,prepro_filelist)] #insert conditions as well.
  file2=prepro_filelist[grepl(f2,prepro_filelist) & grepl(co,prepro_filelist)]
  
  d1=read.csv(file1) #Load the recreated filenames
  
  d2=read.csv(file2) 
  
  #Ensure both dataframes have the same length
  if (nrow(d1)>nrow(d2)){
    d1=d1[1:nrow(d2),]
  } else {
    d2=d2[1:nrow(d1),]
  }
  
  #creation of surrogate files. (Riccardo's idea of having fun and completely overdoing it ;-) ) 
  #per each 2 pairs there are 4 possible surrogate pairs: 
  #pair1person1 w pair2person2; 
  d3=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp2,HR1=d1$HR1,HR2=d2$HR2,Resp1S=d1$Resp1S,Resp2S=d2$Resp2S,HR1S=d1$HR1S,HR2S=d2$HR2S)
  #pair1person2 w pair2person2;
  d4=data.frame(Resp1=d2$Resp1,Resp2=d1$Resp2,HR1=d2$HR1,HR2=d1$HR2,Resp1S=d2$Resp1S,Resp2S=d1$Resp2S,HR1S=d2$HR1S,HR2S=d1$HR2S)
  
  #pair1person1 w pair2person1;
  d5=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp1,HR1=d1$HR1,HR2=d2$HR1,Resp1S=d1$Resp1S,Resp2S=d2$Resp1S,HR1S=d1$HR1S,HR2S=d2$HR1S)
  #pair1person2 w pair2person2
  d6=data.frame(Resp1=d2$Resp2,Resp2=d1$Resp2,HR1=d2$HR2,HR2=d1$HR2,Resp1S=d2$Resp2S,Resp2S=d1$Resp2S,HR1S=d2$HR2S,HR2S=d1$HR2S)
  
  #Make filenames that tells which pairs have been mixed
  file3=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"a_",co,".csv")
  file4=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"b_",co,".csv")
  file5=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"c_",co,".csv")
  file6=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"d_",co,".csv")
  
  #write csv files
  write_csv(d3,file.path("surrogate",file3))
  write_csv(d4,file.path("surrogate",file4))
  write_csv(d5,file.path("surrogate",file5))
  write_csv(d6,file.path("surrogate",file6))
  }
}


#DO THE SAME FOR STUDY 2
prepro_filelist = list.files("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data", pattern = "^pp_Study2.*.csv")

num=str_extract_all(prepro_filelist, "\\d+") #extract all numbers in the file name (i.e Study?_G?_T?)

df= as.data.frame(t(as.data.frame(num))) #create a transposed (i.e. "tipped") dataframe of these number
rownames(df)<-NULL #Remove rownames.

pairs=as.numeric(as.character(unique(df$V2))) #create pairs, i.e. unique group numbers

SurrogateList <- expand.grid(a = pairs, b = pairs) # expand.grid combines all possible combinations of a given vector
SurrogateList = subset(SurrogateList, a != b) # Ensure that pair 1 (i.e group1) doesn't get paired with it self.

setwd("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data")
# Per each condition
for (co in c("Synchronous","TurnTaking","Conversation")){ #loop through a list of condintions 
  for (i in 1:nrow(SurrogateList)){ #and loop through all the pairings of the surrogate list
    
  #start recreate filenames from the working directory based on the matching from the surrogatelist
  f1=paste0('Study2_G',SurrogateList$a[i]) #Insert group numeber from Surrogatelist 
  f2=paste0('Study2_G',SurrogateList$b[i])
  
  file1=prepro_filelist[grepl(f1,prepro_filelist) & grepl(co,prepro_filelist)] #insert conditions as well.
  file2=prepro_filelist[grepl(f2,prepro_filelist) & grepl(co,prepro_filelist)]
  
  d1=read.csv(file1) #Load the recreated filenames
  
  d2=read.csv(file2) 
  
  #Ensure both dataframes have the same length
  if (nrow(d1)>nrow(d2)){
    d1=d1[1:nrow(d2),]
  } else {
    d2=d2[1:nrow(d1),]
  }
  
  #creation of surrogate files. (Riccardo's idea of having fun and completely overdoing it ;-) ) 
  #per each 2 pairs there are 4 possible surrogate pairs: 
  #pair1person1 w pair2person2; 
  d3=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp2,HR1=d1$HR1,HR2=d2$HR2,Resp1S=d1$Resp1S,Resp2S=d2$Resp2S,HR1S=d1$HR1S,HR2S=d2$HR2S)
  #pair1person2 w pair2person2;
  d4=data.frame(Resp1=d2$Resp1,Resp2=d1$Resp2,HR1=d2$HR1,HR2=d1$HR2,Resp1S=d2$Resp1S,Resp2S=d1$Resp2S,HR1S=d2$HR1S,HR2S=d1$HR2S)
  
  #pair1person1 w pair2person1;
  d5=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp1,HR1=d1$HR1,HR2=d2$HR1,Resp1S=d1$Resp1S,Resp2S=d2$Resp1S,HR1S=d1$HR1S,HR2S=d2$HR1S)
  #pair1person2 w pair2person2
  d6=data.frame(Resp1=d2$Resp2,Resp2=d1$Resp2,HR1=d2$HR2,HR2=d1$HR2,Resp1S=d2$Resp2S,Resp2S=d1$Resp2S,HR1S=d2$HR2S,HR2S=d1$HR2S)
  
  #Make filenames that tells which pairs have been mixed
  file3=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"a_",co,".csv")
  file4=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"b_",co,".csv")
  file5=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"c_",co,".csv")
  file6=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"d_",co,".csv")
  
  #write csv files
  write_csv(d3,file.path("surrogate",file3))
  write_csv(d4,file.path("surrogate",file4))
  write_csv(d5,file.path("surrogate",file5))
  write_csv(d6,file.path("surrogate",file6))
  }
}






#Run CRQA analysis on Surrogate data  
HR_Delay = median(crqa_df$HR_Delay, na.rm=TRUE)
HR_Embed = median(crqa_df$HR_Embed, na.rm=TRUE)
HR_Radius = median(crqa_df$HR_Threshold, na.rm=TRUE)+1.2

Resp_Delay = median(crqa_df$Resp_Delay, na.rm=TRUE)
Resp_Embed = median(crqa_df$Resp_Embed, na.rm=TRUE)
Resp_Radius = median(crqa_df$Resp_Threshold, na.rm = TRUE)

HR_Delay
HR_Embed
HR_Radius
Resp_Delay
Resp_Embed
Resp_Radius

study = NULL
group = NULL 
trial = NA 
condition = NULL
HR_RR = NULL 
HR_L = NULL 
HR_DET = NULL 
HR_maxL = NULL 
HR_TT = NULL 
HR_LAM = NULL 
HR_ENTR = NULL
Resp_RR = NULL 
Resp_L = NULL 
Resp_DET = NULL 
Resp_maxL = NULL 
Resp_TT = NULL 
Resp_LAM = NULL 
Resp_ENTR = NULL
Type= NULL 



surrogate_list = list.files("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data/surrogate", pattern = ".*.csv")

setwd("~/Documents/Experimental Methods III/Assignment_4/Data/prepro_data/surrogate")
n=1
for (i in surrogate_list){
  ts = read.delim(i, header=TRUE, sep=",")
  
  #Extract study information 
  study[n] = str_extract(i, regex("\\d{1}"))
  #Extract group information 
  group[n] = str_extract(i, regex("\\d{1,2}&\\d{1,2}[a-z]{1}"))
  #Extract condition information 
  condition[n] = str_extract(i, regex("Synchronous|TurnTaking|Conversation"))
  
  Type[n] = "surrogate"
  
  crqa_HR = try(crqa(ts$HR1S, ts$HR2S, delay=HR_Delay, embed=HR_Embed, radius=HR_Radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  
  HR_RR[n] = crqa_HR$RR
  HR_L[n] = crqa_HR$L
  HR_DET[n] = crqa_HR$DET
  HR_maxL[n] = crqa_HR$maxL
  HR_TT[n] = crqa_HR$TT
  HR_LAM[n] = crqa_HR$LAM
  HR_ENTR[n] = crqa_HR$ENTR
  
  crqa_Resp = try(crqa(ts$Resp1S, ts$Resp2S, delay=Resp_Delay, embed=Resp_Embed, radius=Resp_Radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  
  
  Resp_RR[n] = crqa_Resp$RR
  Resp_L[n] = crqa_Resp$L
  Resp_DET[n] = crqa_Resp$DET
  Resp_maxL[n] = crqa_Resp$maxL
  Resp_TT[n] = crqa_Resp$TT
  Resp_LAM[n] = crqa_Resp$LAM
  Resp_ENTR[n] = crqa_Resp$ENTR
  n=n+1
  }

surro_crqa_data = data.frame(study, trial, group, condition,HR_RR, HR_L, HR_DET, HR_maxL, HR_TT, HR_LAM, HR_ENTR, Resp_RR, Resp_L, Resp_DET, Resp_maxL, Resp_TT, Resp_LAM, Resp_ENTR, Type)

surro_crqa_data = select(surro_crqa_data, study, trial, group, condition,HR_RR, HR_L, HR_DET, HR_maxL, HR_TT, HR_LAM, HR_ENTR, Resp_RR, Resp_L, Resp_DET, Resp_maxL, Resp_TT, Resp_LAM, Resp_ENTR, Type)

#Save the dataframe of all crqa measures for each file (in one df)
write.csv(surro_crqa_data, file="surro_crqa_data")

#Replace all "NA" values to "Conversation"
library(plyr)
revalue(surro_crqa_data$condition, c(NA="Conversation"))
```



Bind real, shuffled and surrogate crqa analysis results in one df
```{r}
all_crqa_data = rbind(crqa_data, shuff_crqa_data, surro_crqa_data)

#In surrogate data, "Conversation" levels in condition had value "NA"
#Replace all NAs with "Conversation"
all_crqa_data[is.na(all_crqa_data)] <- "Conversation"
surro_crqa_data[is.na(surro_crqa_data)] = "Conversation"
#Save the df 
write.csv(all_crqa_data, file="all_crqa_data")

#Read the final df
all_crqa_data = read.csv("~/Documents/Experimental Methods III/Assignment_4/all_crqa_data")




```
2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

There's gonna be coordination between the signals, jsut because of the fact htat we are all humans with a body and a respiration system
- Take the time-series, shuffte it - is the shuffled baseline significantly different from the real time-series? - If not, no actual correlation - if different - its not jsut because our bodies are similar, there is something more

Also test by creating surrogate pairs - people are engaged in the same activity, but not in the same pair - check if it is just the activity that creates coordination, or if it is the task of working in a pair. - The coordination is not just driven by the activity. 

```{r}
library(lme4)
install.packages("lmerTest")
library(lmerTest)

#See means of heart rate RR for each type 
dplyr::group_by(all_crqa_data, Type) %>% 
  dplyr::summarise(mean=mean(HR_RR))

#See means of respiration RR for each type 
dplyr::group_by(all_crqa_data, Type) %>% 
  dplyr::summarise(mean=mean(Resp_RR))

#HEART RATE - recurrence rate predicted by type 
RR_HR_type = lmer(HR_RR~Type+(1|group)+(1|study), all_crqa_data)
summary(RR_HR_type)
#Real estimate = 13.93, SE = 1.19, p<.001 ----- Remember this is the intercept, do not report these numbers! 
#Shuffled estimate = -14.10, SE = 1.06, p<.001
#Surrogate estimate = -0.68, SE = 0.46, p=.56

#RESPIRATION - recurrence rate predicted by type 
RR_Resp_type = lmer(Resp_RR~Type+(1|group)+(1|study), all_crqa_data)
summary(RR_Resp_type)
#Real estiamte = 6.17, SE = 1.12, p<.01.
#Shuffled estimate = -1.07, SE = 1.09, p=.33
#Surrogate estimate = -0.46, SE = 0.80, p=.57


```
3) Do you observe differences in coordination between conditions? Report the models and results.
- (Heart rate RC ~ condition, subset(data, type="real")) - only tests if they are different
- We want to see if the real difference in baseline (surrogate) and real pairs, is different between conditions - thus, we look at the coordination after we removed the coordination due to activity. - So we see if it is only due to the fact that they are doing something together (not just the same activity)
  - (heart rate RC ~ condition*type, data)
- Report results of only one of the models 
```{r}
RR_HR_real = lmer(HR_RR~condition+(1|group)+(1|study), crqa_data)
summary(RR_HR_real)
#Conversation estimate = 11.34, SE = 2.14, p<.05
#Synchrounous estimate = 3.97, SE = 1.60, p<.05
#TurnTaking estimate = 1.97, SE = 1.60, p=.24

RR_HR_type_cond = lmer(HR_RR~condition*Type+(1|group)+(1|study), all_crqa_data)
summary(RR_HR_type_cond)


RR_Resp_real = lmer(Resp_RR~condition+(1|group)+(1|study), crqa_data)
summary(RR_Resp_real)
#Conversation estimate = 6.02, SE = 2.18, p=.08
#Synchronous estimate = 0.30, SE = 1.98, p=.88
#TurnTaking estimate = 0.50, SE = 1.98, p=.80

RR_Resp_type_cond = lmer(Resp_RR~condition*Type+(1|group)+(1|study), all_crqa_data)
summary(RR_Resp_type_cond)


```
### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them
```{r}
#Maybe HR_RR predicted by Resp1S or Resp_RR?
```



2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.
- There's gonna be coordination between the signals, jsut because of the fact htat we are all humans with a body and a respiration system
- Take the time-series, shuffte it - is the shuffled baseline significantly different from the real time-series? - If not, no actual correlation - if different - its not jsut because our bodies are similar, there is something more- Use the function sample()
- Make Loop that shuffles all the time-series 
- Run the entire crqa anlaysis again on these shuffled pairs 
- store in the same df

- Also test by creating surrogate pairs - people are engaged in the same activity, but not in the same pair - check if it is just the activity that creates coordination, or if it is the task of working in a pair. - The coordination is not just driven by the activity. 
  - Just choose one of the possible surrogate pairs for each real pair
  - So for each pair we will have the type "real", "shuffle", and "surrogate" in the dataframe
- Run the entire crqa anlaysis again on these surrogate pairs 
- store in the same df



3) Do you observe differences in coordination between conditions? Report the models and results.
- (Heart rate RC ~ condition, subset(data, type="real")) - only tests if they are different
- We want to see if the real difference in baseline (surrogate) and real pairs, is different between conditions - thus, we look at the coordination after we removed the coordination due to activity. - So we see if it is only due to the fact that they are doing something together (not just the same activity)
  - (heart rate RC ~ condition*type, data)
- Report results of only one of the models 

4) Is respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.
- ONly a conceptual question 
- Using condition as predictor above, is rather indirect, it captures this it in a rough way - what we really want to know is whether coordination of respiration is driving coordination of heart rate 
- What would you do if you were to test this hypothesis 

N.B. to give you a bit more data I included data from last year (Study1) and from your class (Study2). Note that synchronouns and turn-taking are the same across both studies, but the third condition is different: last year it was self-paced joint reading; this year it was the tv-series conversation. So you might want to exclude the self-paced reading (but, up to you!)

## Step by step suggestions to solve the assignment

### Exploring physiological signals

- Choose one pair (one pair, three conditions)
- Load the logs
- Produce a plot of the participants' respiration signal and a different one of the participants' HR signal (for inspecting whether the data is usable)
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3)
- Can you eye-ball which condition if any displays more physiological coordination?

- Run crqa on heart rate and respiration data (find parameters, run crqa)
- Does this tell you more than just eyeballing the plots?

### Systematically pre-process the data
- Loop through all the files (either with a loop or with a function), check which files should be excluded, if any, and save the pre-processed time-series. Tip: plot and visually inspect the data to figure out which should be excluded.
- Run crqa on all the pre-processed time-series and save the output (don't forget to add columns with study, group, condition and trial). Tip: remember to first assess optimal parameters (dimensions, delay, radius) across all timeseries. Tip: it will often fail, just take whatever parameters you get, select optimal across timeseries parameters and run crqa on all timeseries with those. Tip: double check the rr. When I ran the loop, I got very low rr, so I adjusted the radius until the average of rr across all pairs was approx. 4%.

### Creating controls: shuffled controls
 - loop through all pairs and conditions
 - shuffle the timeseries (take a timeseries and rearrange its values in a random order). Tip check the sample() function
 - run crqa and save the output. NB. which delay, embed, radius parameters should you use?
 - statistically compare the crqa indexes in real and shuffled pairs
 
### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair). Tip: Celine will share a commented script
 - Run crqa on all the surrogate pairs and save the output. NB. which delay, embed, radius parameters should you use?
 - Test whether crqa shows a difference between real and surrogate pairs

### Testing effects of conditions
 - make a (probably underpowered) mixed model testing effects of the different conditions on heart rate and respiration coordination
 - N.B: would it make sense to include surrogate pairs? and if so how? what would that tell you?

### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them